Parallel Computing in Tikhonov Regularization Method for Solving the Inverse Problem of Chemical Kinetics

### 2023-03-19 2023/PAVT_Ufa/PAVT2023.tex
fix one comment

%
\authorrunning{K. Barkalov et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
%\tocauthor{Ivar Ekeland and Kim~B.~Bruce, and Roger Temam}
%
\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia \\
	\email{konstantin.barkalov@itmm.unn.ru}
	\and
	Ufa State Petroleum Technological University, Ufa, Russia \\
	\email{leniza.enikeeva@yandex.ru} \\
        Institute of Petrochemistry and Catalysis – Subdivision of the Ufa Federal Research Centre of RAS, Ufa, Russia
}

@@

%Ускорение составило примерно $230$ раз.
The short time for solving the problem when using the parallel algorithm (approximately 14 minutes for 1 problem) allowed a thorough investigation of the dependence of the solution on the regularisation parameter $\alpha$. The table \ref{table_1} shows the minimum values of the objective function found for the corresponding value of the regularization parameter, the time for solving the problem (in seconds) and the time speedup in comparison to the estimated problem solution time when using the sequential method. The method parameter $r=4.0$ from (\ref{R}) and the accuracy $\epsilon = 10^{-4}\left\|b-a\right\|$ were used in the runs. After a given number of iterations or after reaching the specified accuracy with the global search method, the solution was refined by the Hooke-Jeeves local method \cite{HookJeeves} with an accuracy of $\epsilon = 10^{-4}\left\|b-a\right\|$.

@@

In the course of the research, the functional $f(z)$ was calculated without taking into account the addition of the regularization term (figure \ref{fig7}). Then a similar calculation was carried out taking into account the regularization term. The results for $\alpha=0.002$ is $0.527$ – with regularization, and $0.547$, respectively.

@@

@@

%
% ---- Bibliography ----
%
\bibliographystyle{spmpsci}
\bibliography{bibliography}{}

>>>>

%
\authorrunning{K. Barkalov et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
%\tocauthor{Ivar Ekeland and Kim~B.~Bruce, and Roger Temam}
%
\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia \\
	\email{konstantin.barkalov@itmm.unn.ru}
	\and
	Ufa State Petroleum Technological University, Ufa, Russia \\
	\email{leniza.enikeeva@yandex.ru} \\
        \and
        Institute of Petrochemistry and Catalysis – Subdivision of the Ufa Federal Research Centre of RAS, Ufa, Russia
}

@@

%Ускорение составило примерно $230$ раз.
The short time for solving the problem when using the parallel algorithm (approximately 14 minutes for 1 problem) allowed a thorough investigation of the dependence of the solution on the regularisation parameter $\alpha$. Table \ref{table_1} shows the minimum values of the objective function found for the corresponding value of the regularization parameter and the time for solving the problem (in seconds).
%and the time speedup in comparison to the estimated problem solution time when using the sequential method. 
The method parameter $r=4.0$ from (\ref{R}) and the accuracy $\epsilon = 10^{-4}\left\|b-a\right\|$ were used in the runs. After a given number of iterations or after reaching the specified accuracy with the global search method, the solution was refined by the Hooke-Jeeves local method \cite{HookJeeves} with an accuracy of $\epsilon = 10^{-4}\left\|b-a\right\|$.

@@

In the course of the research, the functional $f(z)$ was calculated without taking into account the addition of the regularization term (figure \ref{fig7}). Then a similar calculation was carried out taking into account the regularization term. The results for $\alpha=0.002$ is $0.527$ with regularization, and $0.547$, respectively.

@@

The speedup of the parallel global optimization algorithm was estimated for the problem in which the best solution was found, i.e. the problem with regularization parameter $\alpha = 0.001$.
Since solving the problem in serial mode would require about two days, the acceleration was evaluated in relation to a parallel launch involving a smaller number of processes.
Table \ref{table_2} shows running time and acceleration of the parallel algorithm using 80 and 160 processes in relation to running time using 40 processes

\begin{table}
\caption{Speedup of the parallel  global optimization algorithm}
\label{table_2}
\begin{center}
\begin{tabular}{ccc}
\hline\noalign{\smallskip}
 $p$     & Time (sec.)  & Speedup \\
\hline\noalign{\smallskip}
40	&	2778.7		&	---	\\
80	&	1495.5		&	1.9	\\
160	&	853.9	    &	3.3	\\
\noalign{\smallskip}\hline
\end{tabular}\end{center}\end{table}

@@

%This research was partially funded by the Ministry of Science and Higher Education of the Russian Federation, project no. FSWR-2023-0034, by the Research and Education Mathematical Center, project no. 075-02-2022-883 (development of the parallel global optimization algorithm), and by the Ministry of Science and Higher Education of the Russian Federation, project no. FMRS-2022-0078 (investigation of chemical kinetics problem).

%
% ---- Bibliography ----
%
\bibliographystyle{spmpsci}
\bibliography{example-bibtex}{}
